---
title: "Lab 3"
author: "Sean Kim"
format:
  html:
    embed-resources: true
---

## Question 1:

```{r}
download.file(
  "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
  destfile = file.path("~", "Downloads", "met_all.gz"),
  method   = "libcurl",
  timeout  = 60
)
met <- data.table::fread(file.path("~", "Downloads", "met_all.gz"))

library(tidyverse)
library(data.table)
```

# Question 2: How many columns, rows are there?

```{r}
dim(met)
head(met)
tail(met)
```

There are 2,377,343 rows and 30 columns in the met dataset.

# Question 3: Looking at the variables

```{r}
str(met)
```

# Question 4: Closer look at key variables:

```{r}
table(met$year)
table(met$day)
table(met$hour)
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

```{r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```

Highest weather station is 4113 meters, after replacing 9999.0 values with NA.

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)

```

# Question 5 - Check data against an external data source

```{r}
met <- met[temp>-15]
met2 <- met[order(temp)]
head(met2)
```

All temperatures colder than -15C were removed. Minimum temp -3C in the new dataset.

```{r}
max(met$elev)
min(met$elev)
```

# 6. Summary Statistics

```{r}
elev <- met[elev==max(elev), ]
summary(elev)
```

```{r}
cor(elev$temp, elev$wind.sp, use="complete")
cor(elev$temp, elev$hour, use="complete")
cor(elev$wind.sp, elev$day, use="complete")
cor(elev$wind.sp, elev$hour, use="complete")
cor(elev$temp, elev$day, use="complete")
```

# Question 7: Exploratory Graphs

```{r}
hist(met$elev, breaks=100)
hist(met$temp)
hist(met$wind.sp)

```

```{r}

```

library(leaflet)
leaflet(elev) %\>%

addProviderTiles('OpenStreetMap') %\>% 

addCircles(lat=\~lat,lng=\~lon, opacity=1, fillOpacity=1, radius=100)

```{r}
library(lubridate)
elev$date <- with(elev, ymd_h(paste(year, month, day, hour, sep= ' ')))
summary(elev$date)
elev <- elev[order(date)]
head(elev)
plot(elev$date, elev$temp, type='l')
plot(elev$date, elev$wind.sp, type='l')
```

For temp vs date, there appears to be variation within a day, likely reflecting the temperature differences between day and night, and overall temp seems range between 2-14C and remained similar in that trend throughout the month.

For wind vs date, there are two spikes in wind speed around Aug 18 and Aug 26.

# Question 8: What questions about the data?

Question: What is the relationship between elevation and wind speed in August?

```{r}
summary(met$wind.sp)
met3 <- met[met$wind.sp < 36, ]
dim(met3)
met4 <- met3[order(met3$wind.sp)]
head(met4)
tail(met4)



```

```{r}
met4_daily <- met4[, .(
  wind.sp = mean(wind.sp),
  lat  = mean(lat),
  lon  = mean(lon),
  elev = mean(elev)
  ), by = c("USAFID", "day")][order(wind.sp)]
head(met4_daily)
tail(met4_daily)
```

```{r}
met4_daily[, plot(elev, wind.sp, pch = 19, cex =0.5)]
mod <- lm(wind.sp ~ elev, data = met4_daily)
met4_daily[, plot(
  elev, wind.sp, pch=19, cex=0.5, 
  main = "Wind Speed and Elevation", 
  xlab = "Elevation", ylab = "Wind Speed (mph)")
  ]
abline(mod, lwd=2, col="red")
```

Average wind speed increases with increasing elevation.
